### 2016-06-08 paper/main.tex
update paper

\def\df{df}
\def\FDR{\text{FDR}}
\def\fdr{\text{\lfdr}}
\def\FSR{\text{FSR}}
\def\fsr{\text{lfsr}}

@@

Let $\beta=(\beta_1,\dots,\beta_J)$ denote $J$ ``effects" of interest. For example, in a genomics application 
$\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
that is, testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
more general problem of estimating, and assessing uncertainty in, $\beta_j$.

Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
which by Bayes theorem can be written as 
\begin{equation}
p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
\end{equation}

For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
A simple way to implement the UA is to assume that $g$ is
a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
\begin{align} \label{eqn:beta}
p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
\end{align}
where $\delta_0(\cdot)$ denotes a point mass on 0, and $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyper-parameters, which are non-negative and sum to one, and are to be estimated,
while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
We encourage the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
although of course in practice we use a finite grid -- see Implementation Details.

@@

This simple model features both the key ideas we want to emphasize in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
However, the model also has several additional assumptions that can be relaxed. Specifically,
\begin{enumerate}
\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) implies that $g$ is symmetric about 0.
More flexibility can be obtained using mixtures of uniforms [see equation (\ref{eqn:g-general})]; indeed this 
can allow $g$ to approximate {\it any} unimodal distribution. 
\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
This can be relaxed [see (\ref{eqn:beta-alpha})].
\item The normal likelihood (\ref{eqn:normlik}) can be generalized to a $t$ likelihood [see (\ref{eqn:tlik})].
\end{enumerate}
These embellishments, detailed in Detailed Methods, are all implemented in our software.
Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions
(which are also made by most existing approaches). Correlations among tests certainly arise in practice, either due to genuine correlations 
or due to unmeasured confounders, and their potential impact on estimated FDRs is important to consider
whatever analysis methods are used \cite{efron2007correlation,leek:2007}.

@@

Together, (\ref{eqn:beta})-(\ref{eqn:normlik}) imply that $\bhat_j$ are independent observations from a mixture of normal distributions:
\begin{equation} \label{eqn:marginallik}
p(\bhat | \shat, \pi) =   \prod_j  \left[\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2) \right],
\end{equation}
where we define $\sigma_0:=0$.

The usual EB approach to fitting this model would involve two simple steps:
\begin{enumerate}
\item Estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | \bhat, \shat,\hat{\pi})$.
\end{enumerate}
Both steps 1 and 2 are very straightforward. Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}. (Alternatively a simple EM algorithm can also work well, particularly for modest $J$, say $J<2000$;
see \url{http:://stephenslab.github.io/ash/analysis/IPvsEM.html}.) 
And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
(The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)

@@

As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
In other words it is the posterior probability that $\beta_j$ is actually zero:
\begin{equation} \label{eqn:fdr}
\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
\end{equation}

The $\lfdr$, like most other measures of significance, is rooted in the hypothesis testing paradigm which focuses on 
whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey \cite{tukey1991philosophy} argued that
``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$. Thus asking `Are the effects different?' is foolish." Instead, Tukey (\cite{tukey1962future}, p32) suggested  that one should address
\begin{quote}
...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
\end{quote}
Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.

@@

As our notation suggests, $\lfsr_j$ is analogous to $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
Of course, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
(this follows from the definition because both the events $\beta_j \geq 0$
and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
$\lfsr$ is more robust to modeling assumptions.

From these ``local" measures of significance, we can also compute average error rates over subsets of observations
$\Gamma \subset \{1,\dots,J\}$. For example, 
\begin{equation} \label{eqn:FSRhat}
\widehat\FSR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \lfsr_j 
\end{equation}
is an estimate of the total proportion of errors made if we were to estimate the sign of all effects in $\Gamma$
(an error measure analogous to the usual (tail) FDR).
And, we can define the ``$s$-value" 
\begin{equation} \label{eqn:qvalue}
s_j := \widehat \FSR(\{k: \lfsr_k \leq \lfsr_j\}) 
\end{equation}
as a measure of significance analogous to Storey's $q$-value \cite{storey.03}.
(Replacing $\lfsr_j$ with $\lfdr_j$ in these definitions gives an estimate for the usual $\FDR(\Gamma)$, and an estimated $q$ value for each observation respectively.)

@@

Among previous methods that explicitly consider the FDR and related quantities,
our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
 (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
  %Here we highlight the key differences between our approach and these three existing approaches.

There are two key differences between our approach and all of these three existing methods. First, 
whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
for individual null hypotheses in a similar way; see also \cite{efron1993bayes}.
Using two numbers instead of one clearly has the potential to be more informative, and this is illustrated in Results later (Figure \ref{fig:goodpoor}).

@@

In further support of the UA, note that large-scale regression methods almost always make the UA for the regression coefficients,
which are analogous to the ``effects" we estimate here. 
Common choices of uni-modal distribution for regression coefficients 
include the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
These are all less flexible than our approach, which provides for general uni-modal distributions, and 
it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
Additionally, the UA can be motivated by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).

@@

There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
 Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
Also, they provide no software implementation.

More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic". Some pointers to recent papers in which other relevant
citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}. Much of the literature focusses on the homoskedastic case (i.e. $\shat_j$ all equal) whereas we allow for heteroskedasticity. And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential. Several recent papers consider more flexible non-parametric assumptions on 
$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
estimate (NPMLE) for $g$. These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive. However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
may perform well for point estimation, it may not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
Indeed, one way of thinking about the UA is as a way to regularize $g$.

@@

Here we consider the effects of making the UA. 
To isolate these effects we consider the simplest case, where every observation has the same
standard error, $s_j=1$ and all methods are provided that information. That is,
$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$. In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the primary difference among methods lies in how they estimate $g$.

To briefly summarize the results in this section:
\begin{enumerate}
\item The UA can produce quite different results from existing methods.
\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
\end{enumerate}

@@

\begin{figure}
\center \includegraphics[height=6in]{../analysis/figure/makefig_FDReg.Rmd/decomp_ZA-1.pdf}
\caption{Illustration that the unimodal assumption (UA) in \ash can produce very different results from existing methods.
The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan). In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
The three existing methods (\qvalue, \locfdr, \mixfdr) all produce a ``hole" in the alternative $z$ score distribution around 0.
In contrast the method introduced here (\ash) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition. (In this case the \ash decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
\end{figure}

We illustrate the effects of the UA, and the difference with existing methods, using a simple simulated dataset
 with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
\qvalue, \locfdr, \mixfdr and \ash to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
Here we are using the fact that these methods all provide an estimate of the \lfdr{} for each observation, which
implies such a decomposition;  specifically the average \lfdr{} within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component. The results (Figure \ref{fig:ZA}) illustrate a clear difference between the existing methods and our method:
all the existing methods have a ``hole"  in the alternative $z$ score distribution near 0, whereas \ash, due to the UA, has a mode near 0.
%Although \mixfdr does not formally make the \sZA, its decomposition
%exhibits a similar hole. 
%In contrast, due to the UA, the alternative $z$ score distribution for \ash is required to have a mode at 0, 
%effectively ``filling in" the hole.  
(Of course the null distribution also has a peak at 0, and the local fdr 
 under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)

This qualitative difference among methods in the inferred alternative distribution of the $z$ scores is quite general, and 
also occurs in simulations where most effects are null (e.g.~\url{http://stephenslab.github.io/ash/analysis/referee_uaza.html}). To understand why the alternative distribution of $z$ scores from $\locfdr$ and $\qvalue$ has a hole at zero, note that neither
of these methods explicitly models the alternative distribution: instead they simply subtract a null distribution (of $z$ scores or $p$ values) from the observed empirical distribution, letting the alternative
distribution be defined implicitly, by what remains.
 In deciding how much null distribution to subtract -- that is, in estimating the null proportion, $\pi_0$ -- 
 both methods assume that all $z$ scores near zero (or, equivalently,
all $p$ values near 1) are null. The consequence of this is that their (implicitly-defined) distribution for the alternative $z$ scores has 
a ``hole" at 0 -- quite different from our assumption of a mode at zero. (Why \mixfdr exhibits similar behaviour is less clear, since it does explicitly
model the alternative distribution; however we believe it may be due to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}.)

Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA. Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ash's decomposition of $z$ scores, and imagine increasing $\pi_0$ further. This would increase the null component (dark blue) at the expense of
the alternative component (light blue). Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0. The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears. In contrast the 
existing methods effectively push the dark blue component until the light-blue component {\it disappears} at 0. See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive demonstration.

@@

%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
%
%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
%and so all effects are non-zero and no nulls hold. This scenario is rather different
%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
%``mostly non-null"
%scenario, and we would agree! On the other hand, the distribution of $p$ values
%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
%Thus it seems desirable to have methods that perform well in both cases, and this is
%an important feature of the methods we introduce here.
%
%TODO: put FDRs on graph? Simulate with 50\% null?

>>>>

\def\df{df}
\def\FDR{\text{FDR}}
\def\FSR{\text{FSR}}
\def\fsr{\text{lfsr}}

@@

Let $\beta=(\beta_1,\dots,\beta_J)$ denote $J$ ``effects" of interest. For example, in a genomics application 
$\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
We tackle both the problem of testing the null hypotheses $H_j:\beta_j=0$, and the  
more general problem of estimating, and assessing uncertainty in, $\beta_j$.

Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given $\bhat,\shat$,
which by Bayes theorem is
\begin{equation}
p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
\end{equation}

For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
A simple way to implement the UA is to assume that $g$ is
a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
\begin{align} \label{eqn:beta}
p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
\end{align}
where $\delta_0(\cdot)$ denotes a point mass on 0, and $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
Here we take $\sigma_1,\dots,\sigma_K$ to be a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
We encourage the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
although of course in practice we use a finite grid -- see Implementation Details.
The mixture proportions $\pi=(\pi_0,\dots,\pi_K)$, which are non-negative and sum to one, are 
hyper-parameters to be estimated.

@@

This simple model features both the key ideas we want to emphasize in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
However, the model also has several additional assumptions that can be relaxed. Specifically,
\begin{enumerate}
\item  The form (\ref{eqn:mixnorm}) implies that $g$ is symmetric about 0.
More flexibility can be obtained using mixtures of uniforms [Equation (\ref{eqn:g-general})]; indeed this 
allows $g$ to approximate {\it any} unimodal distribution. 
\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
This can be relaxed [see (\ref{eqn:beta-alpha})].
\item The normal likelihood (\ref{eqn:normlik}) can be generalized to a $t$ likelihood [see (\ref{eqn:tlik})].
\end{enumerate}
These embellishments, detailed in Detailed Methods, are implemented in our software.
Other limitations are harder to relax, most notably the independence and conditional independence assumptions
(which are also made by most existing approaches). Correlations among tests certainly arise in practice, either due to genuine correlations 
or due to unmeasured confounders, and their potential impact on estimated FDRs is important to consider
whatever analysis methods are used \cite{efron2007correlation,leek:2007}.

@@

Together, (\ref{eqn:beta})-(\ref{eqn:normlik}) imply that $\bhat_j$ are independent with
\begin{equation} \label{eqn:marginallik}
p(\bhat | \shat, \pi) =   \prod_j  \left[\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2) \right],
\end{equation}
where we define $\sigma_0:=0$.

The usual EB approach to fitting this model would involve two steps:
\begin{enumerate}
\item Estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | \bhat, \shat,\hat{\pi})$.
\end{enumerate}
Both steps are straightforward. Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}. (Alternatively a simple EM algorithm can also work well, particularly for modest $J$; see \url{http:://stephenslab.github.io/ash/analysis/IPvsEM.html}.) 
And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$. This simple device may be useful in other applications.

@@

As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}
\begin{equation} \label{eqn:fdr}
\lfdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
\end{equation}
In words, $\lfdr_j$ is the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.

The $\lfdr$, like most measures of significance, is rooted in the hypothesis testing paradigm which focuses on 
whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey \cite{tukey1991philosophy} argued that
``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$. Thus asking `Are the effects different?' is foolish." Instead, Tukey (\cite{tukey1962future}, p32) suggested  that one should address
\begin{quote}
...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
\end{quote}
Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.

@@

As our notation suggests, $\lfsr_j$ is analogous to $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
Of course, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
(this follows from the definition because both the events $\beta_j \geq 0$
and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
In this sense $\lfsr$ is a more conservative measure of significance than $\lfdr$. 
More importantly, $\lfsr$ is more robust to modeling assumptions (see Results).

From these ``local" measures of significance, we can also compute average error rates over subsets of observations
$\Gamma \subset \{1,\dots,J\}$. For example, 
\begin{equation} \label{eqn:FSRhat}
\widehat\FSR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \lfsr_j 
\end{equation}
is an estimate of the total proportion of errors made if we were to estimate the sign of all effects in $\Gamma$
(an error measure analogous to the usual (tail) FDR).
And, we can define the ``$s$-value" 
\begin{equation} \label{eqn:svalue}
s_j := \widehat \FSR(\{k: \lfsr_k \leq \lfsr_j\}) 
\end{equation}
as a measure of significance analogous to Storey's $q$-value \cite{storey.03}.
(Replacing $\lfsr_j$ with $\lfdr_j$ in (\ref{eqn:FSRhat}) and (\ref{eqn:svalue}) gives estimates for the usual $\FDR(\Gamma)$, and the $q$-values respectively.)

@@

Among previous methods that explicitly consider FDR,
our work seems naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
 (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
  %Here we highlight the key differences between our approach and these three existing approaches.

There are two key differences between our approach and these existing methods. First, 
whereas these existing methods take as input a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- for each effect, 
we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayes Factors testing
individual null hypotheses in a similar way; see also \cite{efron1993bayes}.
Using two numbers instead of one clearly has the potential to be more informative, as illustrated in Results 
(Figure \ref{fig:goodpoor}).

@@

In further support of the UA, note that large-scale regression methods almost always make the UA for the regression coefficients,
which are analogous to the ``effects" we estimate here. 
Common choices of unimodal distribution for regression coefficients 
include the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
These are all less flexible than our approach, which provides for general unimodal distributions, and 
it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
Additionally, the UA can be motivated by its effect on point estimates, which is to ``shrink" the estimates towards the mode - such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
make parametric assumptions that obey the UA (e.g.~\cite{johnstone2004needles}).

@@

There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
 Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
However, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
Also, they provide no software implementation.

More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical Bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic". Some pointers to recent papers in which other relevant
citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex, efron2016empirical}. 
Much of the literature focusses on the homoskedastic case (i.e.~$\shat_j$ all equal) whereas we allow for heteroskedasticity. And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential. Several recent papers consider more flexible non-parametric assumptions on 
$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
estimate (NPMLE) for $g$. These methods may be useful in settings where the UA assumption is considered too restrictive. However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
may perform well for point estimation, it may not adequately reflect uncertainty in $\beta_j$. To address this 
some regularization on $g$ (e.g.~as in \cite{efron2016empirical}) may be necessary.
Indeed, one way of thinking about the UA is as a way to regularize $g$.

@@

Here we consider the effects of making the UA. 
To isolate these effects we consider the simplest case, where every observation has the same
standard error, $s_j=1$. That is,
$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$. In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$. Thus the primary difference among methods in this setting is that \ash makes the UA and other methods do not.

To briefly summarize the results in this section:
\begin{enumerate}
\item The UA can produce quite different results from existing methods.
\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
\item The UA yields a procedure that is numerically and statistically stable, and is somewhat robust to deviations from unimodality.
\end{enumerate}

@@

\begin{figure}
\center \includegraphics[height=6in]{../analysis/figure/makefig_FDReg.Rmd/decomp_ZA-1.pdf}
\caption{Illustration that the unimodal assumption (UA) in \ash can produce very different results from existing methods.
The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan). In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
The three existing methods (\qvalue, \locfdr, \mixfdr) all produce a ``hole" in the alternative $z$ score distribution around 0.
In contrast \ash makes the UA -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition. (In this case the \ash decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
\end{figure}

We illustrate the effects of the UA with a simple simulation, 
 with effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Though no effects are null, there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used
\qvalue, \locfdr, \mixfdr and \ash to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
Here we are using the fact that these methods all provide an estimated $\lfdr_j$ for each observation $j$, which
implies such a decomposition;  specifically the average $\lfdr_j$ within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component. The results (Figure \ref{fig:ZA}) illustrate a clear difference between \ash and existing methods:
the existing methods have a ``hole"  in the alternative $z$ score distribution near 0, whereas \ash, due to the UA, has a mode near 0.
%Although \mixfdr does not formally make the \sZA, its decomposition
%exhibits a similar hole. 
%In contrast, due to the UA, the alternative $z$ score distribution for \ash is required to have a mode at 0, 
%effectively ``filling in" the hole.  
(Of course the null distribution also has a peak at 0, and the \lfdr{} 
 under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)

This qualitative difference among methods is quite general, and 
also occurs in simulations where most effects are null (e.g.~\url{http://stephenslab.github.io/ash/analysis/referee_uaza.html}). To understand why the alternative distribution of $z$ scores from $\locfdr$ and $\qvalue$ has a hole at zero, note that neither
of these methods explicitly models the alternative distribution: instead they simply subtract a null distribution (of $z$ scores or $p$ values) from the observed empirical distribution, letting the alternative
distribution be defined implicitly, by what remains.
 In deciding how much null distribution to subtract -- that is, in estimating the null proportion, $\pi_0$ -- 
 both methods assume that all $z$ scores near zero (or, equivalently,
all $p$ values near 1) are null. The consequence of this is that their (implicitly-defined) distribution for the alternative $z$ scores has 
a ``hole" at 0 -- quite different from our assumption of a mode at zero. (Why \mixfdr exhibits similar behaviour is less clear, since it does explicitly
model the alternative distribution; however we believe it may be due to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}.)

Figure \ref{fig:ZA} is also helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA. Specifically, consider the panel that shows \ash's decomposition of $z$ scores, and imagine increasing $\pi_0$ further. This would increase the null component (dark blue) at the expense of
the alternative component (light blue). Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0. The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears. In contrast the 
existing methods effectively push the dark blue component until the light-blue component {\it disappears} at 0. See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive demonstration.

@@
